---
title: "Codebook"
author: "Ekaterina Kravtchenko"
date: "8/15/2018"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(intsvy)
library(DescTools)
library(scales)
```

## Loading data

First, I will load the dictionary for easy reference.

```{r dictionary, cache=TRUE, message=FALSE, warning=FALSE}
dict <- read_csv("pisadict2012.csv") %>% spread(X1, x)
dict[["AGE"]]
```

Now I will load the data, and take a preliminary look.  I will also make a copy of the data to clean.  The data file itself is not included in the repository, as it is too large, but it is identical to the file supplied by Udacity.

```{r load data, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
pisa <- pisa_orig <- fread("pisa2012.csv")
```

```{r glimpse}
glimpse(pisa_orig)
```

Here I remove any white space padding.

```{r trim white space, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
pisa <- mutate_if(pisa, is.character, trimws) %>% 
  mutate_if(is.character, as.factor)
```

## Regional data

I get the regional data up here, as the columns are more difficult to manipulate once they are renamed to be more descriptive.

I use the `intsvy` `R` package with `pisa` configuration to replicate country-level scores and compute regional scores.

### School

I'll first see if there's enough data per school to make reasonable generalizations:

```{r data per school, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
pisa %>% 
  group_by(STRATUM,SCHOOLID) %>% 
  tally() %>%
  arrange(desc(n))
```

Many schools have very few records - including records for only one individual student, at which point the data becomes uninterpretable (https://nces.ed.gov/surveys/pisa/faq.asp#10).  This makes school too fine-grained a granularity to use to look at the data.

### Country-level

Below, I replicate country scores for PISA 2012.

#### Math

```{r country math, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
country_math <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MATH","PV2MATH","PV3MATH","PV4MATH","PV5MATH"), config=pisa_conf, by="CNT") %>% select(CNT, Mean) %>% rename(`Math (country)` = Mean)

head(arrange(country_math, desc(`Math (country)`)))
```

##### Math Subscale: Change and Relationships

```{r country math change, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
country_math_cr <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MACC","PV2MACC","PV3MACC","PV4MACC","PV5MACC"), config=pisa_conf, by="CNT") %>% select(CNT, Mean) %>% rename(`Math.ChangeRel (country)` = Mean)

head(arrange(country_math_cr, desc(`Math.ChangeRel (country)`)))
```

##### Math Subscale: Quantity

```{r country math quantity, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
country_math_q <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MACQ","PV2MACQ","PV3MACQ","PV4MACQ","PV5MACQ"), config=pisa_conf, by="CNT") %>% select(CNT, Mean) %>% rename(`Math.Quantity (country)` = Mean)

head(arrange(country_math_q, desc(`Math.Quantity (country)`)))
```

##### Math Subscale: Space and Shape

```{r country math space, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
country_math_ss <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MACS","PV2MACS","PV3MACS","PV4MACS","PV5MACS"), config=pisa_conf, by="CNT") %>% select(CNT, Mean) %>% rename(`Math.SpaceShape (country)` = Mean)

head(arrange(country_math_ss, desc(`Math.SpaceShape (country)`)))
```

##### Math Subscale: Uncertainty and Data

```{r country math uncertainty, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
country_math_uncd <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MACU","PV2MACU","PV3MACU","PV4MACU","PV5MACU"), config=pisa_conf, by="CNT") %>% select(CNT, Mean) %>% rename(`Math.UncData (country)` = Mean)

head(arrange(country_math_uncd, desc(`Math.UncData (country)`)))
```

##### Math Subscale: Employ

```{r country math employ, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
country_math_emp <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MAPE","PV2MAPE","PV3MAPE","PV4MAPE","PV5MAPE"), config=pisa_conf, by="CNT") %>% select(CNT, Mean) %>% rename(`Mean.Employ (country)` = Mean)

head(arrange(country_math_emp, desc(`Mean.Employ (country)`)))
```

##### Math Subscale: Formulate

```{r country math formulate, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
country_math_form <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MAPF","PV2MAPF","PV3MAPF","PV4MAPF","PV5MAPF"), config=pisa_conf, by="CNT") %>% select(CNT, Mean) %>% rename(`Math.Formulate (country)` = Mean)

head(arrange(country_math_form, desc(`Math.Formulate (country)`)))
```

##### Math Subscale: Interpret

```{r country math interpret, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
country_math_int <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MAPI","PV2MAPI","PV3MAPI","PV4MAPI","PV5MAPI"), config=pisa_conf, by="CNT") %>% select(CNT, Mean) %>% rename(`Math.Interpret (country)` = Mean)

head(arrange(country_math_int, desc(`Math.Interpret (country)`)))
```

#### Reading

```{r country reading, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
country_read <- intsvy.mean.pv(data = pisa, pvnames=c("PV1READ","PV2READ","PV3READ","PV4READ","PV5READ"), config=pisa_conf, by="CNT") %>% select(CNT, Mean) %>% rename(`Reading (country)` = Mean)

head(arrange(country_read, desc(`Reading (country)`)))
```

#### Science

```{r country science, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
country_sci <- intsvy.mean.pv(data = pisa, pvnames=c("PV1SCIE","PV2SCIE","PV3SCIE","PV4SCIE","PV5SCIE"), config=pisa_conf, by="CNT") %>% select(CNT, Mean) %>% rename(`Science (country)` = Mean)

head(arrange(country_sci, desc(`Science (country)`)))
```

#### Join data

```{r country join, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
country_data <- list(country_math, country_math_cr, country_math_q, country_math_ss, country_math_uncd, country_math_emp, country_math_form, country_math_int, country_read, country_sci) %>%
  reduce(left_join, by = "CNT")
```

### Sub-region

Below, I compute more fine-grained regional scores.

#### Math

```{r subregion math, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
subregion_math <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MATH","PV2MATH","PV3MATH","PV4MATH","PV5MATH"), config=pisa_conf, by="STRATUM") %>% select(STRATUM, Mean) %>% rename(`Math (region)` = Mean)

head(arrange(subregion_math, desc(`Math (region)`)))
```

##### Math Subscale: Change and Relationships

```{r subregion math change, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
subregion_math_cr <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MACC","PV2MACC","PV3MACC","PV4MACC","PV5MACC"), config=pisa_conf, by="STRATUM") %>% select(STRATUM, Mean) %>% rename(`Math.ChangeRel (region)` = Mean)

head(arrange(subregion_math_cr, desc(`Math.ChangeRel (region)`)))
```

##### Math Subscale: Quantity

```{r subregion math quantity, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
subregion_math_q <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MACQ","PV2MACQ","PV3MACQ","PV4MACQ","PV5MACQ"), config=pisa_conf, by="STRATUM") %>% select(STRATUM, Mean) %>% rename(`Math.Quantity (region)` = Mean)

head(arrange(subregion_math_q, desc(`Math.Quantity (region)`)))
```

##### Math Subscale: Space and Shape

```{r subregion math space, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
subregion_math_ss <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MACS","PV2MACS","PV3MACS","PV4MACS","PV5MACS"), config=pisa_conf, by="STRATUM") %>% select(STRATUM, Mean) %>% rename(`Math.SpaceShape (region)` = Mean)

head(arrange(subregion_math_ss, desc(`Math.SpaceShape (region)`)))
```

##### Math Subscale: Uncertainty and Data

```{r subregion math uncertainty, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
subregion_math_uncd <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MACU","PV2MACU","PV3MACU","PV4MACU","PV5MACU"), config=pisa_conf, by="STRATUM") %>% select(STRATUM, Mean) %>% rename(`Math.UncData (region)` = Mean)

head(arrange(subregion_math_uncd, desc(`Math.UncData (region)`)))
```

##### Math Subscale: Employ

```{r subregion math employ, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
subregion_math_emp <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MAPE","PV2MAPE","PV3MAPE","PV4MAPE","PV5MAPE"), config=pisa_conf, by="STRATUM") %>% select(STRATUM, Mean) %>% rename(`Mean.Employ (region)` = Mean)

head(arrange(subregion_math_emp, desc(`Mean.Employ (region)`)))
```

##### Math Subscale: Formulate

```{r subregion math formulate, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
subregion_math_form <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MAPF","PV2MAPF","PV3MAPF","PV4MAPF","PV5MAPF"), config=pisa_conf, by="STRATUM") %>% select(STRATUM, Mean) %>% rename(`Math.Formulate (region)` = Mean)

head(arrange(subregion_math_form, desc(`Math.Formulate (region)`)))
```

##### Math Subscale: Interpret

```{r subregion math interpret, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
subregion_math_int <- intsvy.mean.pv(data = pisa, pvnames=c("PV1MAPI","PV2MAPI","PV3MAPI","PV4MAPI","PV5MAPI"), config=pisa_conf, by="STRATUM") %>% select(STRATUM, Mean) %>% rename(`Math.Interpret (region)` = Mean)

head(arrange(subregion_math_int, desc(`Math.Interpret (region)`)))
```

#### Reading

```{r subregion reading, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
subregion_read <- intsvy.mean.pv(data = pisa, pvnames=c("PV1READ","PV2READ","PV3READ","PV4READ","PV5READ"), config=pisa_conf, by="STRATUM") %>% select(STRATUM, Mean) %>% rename(`Reading (region)` = Mean)

head(arrange(subregion_read, desc(`Reading (region)`)))
```

#### Science

```{r subregion science, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
subregion_sci <- intsvy.mean.pv(data = pisa, pvnames=c("PV1SCIE","PV2SCIE","PV3SCIE","PV4SCIE","PV5SCIE"), config=pisa_conf, by="STRATUM") %>% select(STRATUM, Mean) %>% rename(`Science (region)` = Mean)

head(arrange(subregion_sci, desc(`Science (region)`)))
```

#### Join data

```{r subregion join, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
subregion_data <- list(subregion_math, subregion_math_cr, subregion_math_q, subregion_math_ss, subregion_math_uncd, subregion_math_emp, subregion_math_form, subregion_math_int, subregion_read, subregion_sci) %>%
  reduce(left_join, by = "STRATUM")
```

## Cleaning

### Dates

`99` likely signifies that the month of birth is not known; in any case, `99` is not a month.

```{r fix month NA, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
pisa[pisa$ST03Q01=="99",]$ST03Q01 <- NA
```

### Rename columns

Here I will rename columns for ease of identifying measures once in Tableau, and selecting which measures to plot.

#### PISA data

```{r rename columns, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
pisa <- pisa %>%
  rename("CNT: Country code 3-character"=CNT,
  "SUBNATIO: Adjudicated sub-region code 7-digit code (3-digit country code + region ID + stratum ID)"=SUBNATIO,
  "STRATUM: Stratum ID 7-character (cnt + region ID + original stratum ID)"=STRATUM,
  "OECD: OECD country"=OECD,
  "NC: National Centre 6-digit Code"=NC,
  "SCHOOLID: School ID 7-digit (region ID + stratum ID + 3-digit school ID)"=SCHOOLID,
  "STIDSTD: Student ID"=STIDSTD,
  "ST01Q01: International Grade"=ST01Q01,
  "ST02Q01: National Study Programme"=ST02Q01,
  "ST03Q01: Birth - Month"=ST03Q01,
  "ST03Q02: Birth - Year"=ST03Q02,
  "ST04Q01: Gender"=ST04Q01,
  "ST05Q01: Attend <ISCED 0>"=ST05Q01,
  "ST06Q01: Age at <ISCED 1>"=ST06Q01,
  "ST07Q01: Repeat - <ISCED 1>"=ST07Q01,
  "ST07Q02: Repeat - <ISCED 2>"=ST07Q02,
  "ST07Q03: Repeat - <ISCED 3>"=ST07Q03,
  "ST08Q01: Truancy - Late for School"=ST08Q01,
  "ST09Q01: Truancy - Skip whole school day"=ST09Q01,
  "ST115Q01: Truancy - Skip classes within school day"=ST115Q01,
  "ST11Q01: At Home - Mother"=ST11Q01,
  "ST11Q02: At Home - Father"=ST11Q02,
  "ST11Q03: At Home - Brothers"=ST11Q03,
  "ST11Q04: At Home - Sisters"=ST11Q04,
  "ST11Q05: At Home - Grandparents"=ST11Q05,
  "ST11Q06: At Home - Others"=ST11Q06,
  "ST13Q01: Mother <Highest Schooling>"=ST13Q01,
  "ST14Q01: Mother Qualifications - <ISCED level 6>"=ST14Q01,
  "ST14Q02: Mother Qualifications - <ISCED level 5A>"=ST14Q02,
  "ST14Q03: Mother Qualifications - <ISCED level 5B>"=ST14Q03,
  "ST14Q04: Mother Qualifications - <ISCED level 4>"=ST14Q04,
  "ST15Q01: Mother Current Job Status"=ST15Q01,
  "ST17Q01: Father <Highest Schooling>"=ST17Q01,
  "ST18Q01: Father Qualifications - <ISCED level 6>"=ST18Q01,
  "ST18Q02: Father Qualifications - <ISCED level 5A>"=ST18Q02,
  "ST18Q03: Father Qualifications - <ISCED level 5B>"=ST18Q03,
  "ST18Q04: Father Qualifications - <ISCED level 4>"=ST18Q04,
  "ST19Q01: Father Current Job Status"=ST19Q01,
  "ST20Q01: Country of Birth International - Self"=ST20Q01,
  "ST20Q02: Country of Birth International - Mother"=ST20Q02,
  "ST20Q03: Country of Birth International - Father"=ST20Q03,
  "ST21Q01: Age of arrival in country of test"=ST21Q01,
  "ST25Q01: International Language at Home"=ST25Q01,
  "ST26Q01: Possessions - desk"=ST26Q01,
  "ST26Q02: Possessions - own room"=ST26Q02,
  "ST26Q03: Possessions - study place"=ST26Q03,
  "ST26Q04: Possessions - computer"=ST26Q04,
  "ST26Q05: Possessions - software"=ST26Q05,
  "ST26Q06: Possessions - Internet"=ST26Q06,
  "ST26Q07: Possessions - literature"=ST26Q07,
  "ST26Q08: Possessions - poetry"=ST26Q08,
  "ST26Q09: Possessions - art"=ST26Q09,
  "ST26Q10: Possessions - textbooks"=ST26Q10,
  "ST26Q11: Possessions - technical reference books"=ST26Q11,
  "ST26Q12: Possessions - dictionary"=ST26Q12,
  "ST26Q13: Possessions - dishwasher"=ST26Q13,
  "ST26Q14: Possessions - DVD"=ST26Q14,
  "ST26Q15: Possessions - Country item 1"=ST26Q15,
  "ST26Q16: Possessions - Country item 2"=ST26Q16,
  "ST26Q17: Possessions - Country item 3"=ST26Q17,
  "ST27Q01: How many - cellular phones"=ST27Q01,
  "ST27Q02: How many - televisions"=ST27Q02,
  "ST27Q03: How many - computers"=ST27Q03,
  "ST27Q04: How many - cars"=ST27Q04,
  "ST27Q05: How many - rooms bath or shower"=ST27Q05,
  "ST28Q01: How many books at home"=ST28Q01,
  "ST29Q01: Math Interest - Enjoy Reading"=ST29Q01,
  "ST29Q02: Instrumental Motivation - Worthwhile for Work"=ST29Q02,
  "ST29Q03: Math Interest - Look Forward to Lessons"=ST29Q03,
  "ST29Q04: Math Interest - Enjoy Maths"=ST29Q04,
  "ST29Q05: Instrumental Motivation - Worthwhile for Career Chances"=ST29Q05,
  "ST29Q06: Math Interest - Interested"=ST29Q06,
  "ST29Q07: Instrumental Motivation - Important for Future Study"=ST29Q07,
  "ST29Q08: Instrumental Motivation - Helps to Get a Job"=ST29Q08,
  "ST35Q01: Subjective Norms - Friends Do Well in Mathematics"=ST35Q01,
  "ST35Q02: Subjective Norms - Friends Work Hard on Mathematics"=ST35Q02,
  "ST35Q03: Subjective Norms - Friends Enjoy Mathematics Tests"=ST35Q03,
  "ST35Q04: Subjective Norms - Parents Believe Studying Mathematics Is Important"=ST35Q04,
  "ST35Q05: Subjective Norms - Parents Believe Mathematics Is Important for Career"=ST35Q05,
  "ST35Q06: Subjective Norms - Parents Like Mathematics"=ST35Q06,
  "ST37Q01: Math Self-Efficacy - Using a Train Timetable"=ST37Q01,
  "ST37Q02: Math Self-Efficacy - Calculating TV Discount"=ST37Q02,
  "ST37Q03: Math Self-Efficacy - Calculating Square Metres of Tiles"=ST37Q03,
  "ST37Q04: Math Self-Efficacy - Understanding Graphs in Newspapers"=ST37Q04,
  "ST37Q05: Math Self-Efficacy - Solving Equation 1"=ST37Q05,
  "ST37Q06: Math Self-Efficacy - Distance to Scale"=ST37Q06,
  "ST37Q07: Math Self-Efficacy - Solving Equation 2"=ST37Q07,
  "ST37Q08: Math Self-Efficacy - Calculate Petrol Consumption Rate"=ST37Q08,
  "ST42Q01: Math Anxiety - Worry That It Will Be Difficult"=ST42Q01,
  "ST42Q02: Math Self-Concept - Not Good at Maths"=ST42Q02,
  "ST42Q03: Math Anxiety - Get Very Tense"=ST42Q03,
  "ST42Q04: Math Self-Concept - Get Good <Grades>"=ST42Q04,
  "ST42Q05: Math Anxiety - Get Very Nervous"=ST42Q05,
  "ST42Q06: Math Self-Concept - Learn Quickly"=ST42Q06,
  "ST42Q07: Math Self-Concept - One of Best Subjects"=ST42Q07,
  "ST42Q08: Math Anxiety - Feel Helpless"=ST42Q08,
  "ST42Q09: Math Self-Concept - Understand Difficult Work"=ST42Q09,
  "ST42Q10: Math Anxiety - Worry About Getting Poor Grades"=ST42Q10,
  "ST43Q01: Perceived Control - Can Succeed with Enough Effort"=ST43Q01,
  "ST43Q02: Perceived Control - Doing Well is Completely Up to Me"=ST43Q02,
  "ST43Q03: Perceived Control - Family Demands and Problems"=ST43Q03,
  "ST43Q04: Perceived Control - Different Teachers"=ST43Q04,
  "ST43Q05: Perceived Control - If I Wanted I Could Perform Well"=ST43Q05,
  "ST43Q06: Perceived Control - Perform Poorly Regardless"=ST43Q06,
  "ST44Q01: Attributions to Failure - Not Good at Maths Problems"=ST44Q01,
  "ST44Q03: Attributions to Failure - Teacher Did Not Explain Well"=ST44Q03,
  "ST44Q04: Attributions to Failure - Bad Guesses"=ST44Q04,
  "ST44Q05: Attributions to Failure - Material Too Hard"=ST44Q05,
  "ST44Q07: Attributions to Failure - Teacher Didnt Get Students Interested"=ST44Q07,
  "ST44Q08: Attributions to Failure - Unlucky"=ST44Q08,
  "ST46Q01: Math Work Ethic - Homework Completed in Time"=ST46Q01,
  "ST46Q02: Math Work Ethic - Work Hard on Homework"=ST46Q02,
  "ST46Q03: Math Work Ethic - Prepared for Exams"=ST46Q03,
  "ST46Q04: Math Work Ethic - Study Hard for Quizzes"=ST46Q04,
  "ST46Q05: Math Work Ethic - Study Until I Understand Everything"=ST46Q05,
  "ST46Q06: Math Work Ethic - Pay Attention in Classes"=ST46Q06,
  "ST46Q07: Math Work Ethic - Listen in Classes"=ST46Q07,
  "ST46Q08: Math Work Ethic - Avoid Distractions When Studying"=ST46Q08,
  "ST46Q09: Math Work Ethic - Keep Work Organized"=ST46Q09,
  "ST48Q01: Math Intentions - Mathematics vs. Language Courses After School"=ST48Q01,
  "ST48Q02: Math Intentions - Mathematics vs. Science Related Major in College"=ST48Q02,
  "ST48Q03: Math Intentions - Study Harder in Mathematics vs. Language Classes"=ST48Q03,
  "ST48Q04: Math Intentions - Take Maximum Number of Mathematics vs. Science Classes"=ST48Q04,
  "ST48Q05: Math Intentions - Pursuing a Career That Involves Mathematics vs. Science"=ST48Q05,
  "ST49Q01: Math Behaviour - Talk about Maths with Friends"=ST49Q01,
  "ST49Q02: Math Behaviour - Help Friends with Maths"=ST49Q02,
  "ST49Q03: Math Behaviour - <Extracurricular> Activity"=ST49Q03,
  "ST49Q04: Math Behaviour - Participate in Competitions"=ST49Q04,
  "ST49Q05: Math Behaviour - Study More Than 2 Extra Hours a Day"=ST49Q05,
  "ST49Q06: Math Behaviour - Play Chess"=ST49Q06,
  "ST49Q07: Math Behaviour - Computer programming"=ST49Q07,
  "ST49Q09: Math Behaviour - Participate in Math Club"=ST49Q09,
  "ST53Q01: Learning Strategies - Important Parts vs. Existing Knowledge vs. Learn by Heart"=ST53Q01,
  "ST53Q02: Learning Strategies - Improve Understanding vs. New Ways vs. Memory"=ST53Q02,
  "ST53Q03: Learning Strategies - Other Subjects vs. Learning Goals vs. Rehearse Problems"=ST53Q03,
  "ST53Q04: Learning Strategies - Repeat Examples vs. Everyday Applications vs. More Information"=ST53Q04,
  "ST55Q01: Out of school lessons - <test lang>"=ST55Q01,
  "ST55Q02: Out of school lessons - <maths>"=ST55Q02,
  "ST55Q03: Out of school lessons - <science>"=ST55Q03,
  "ST55Q04: Out of school lessons - other"=ST55Q04,
  "ST57Q01: Out-of-School Study Time - Homework"=ST57Q01,
  "ST57Q02: Out-of-School Study Time - Guided Homework"=ST57Q02,
  "ST57Q03: Out-of-School Study Time - Personal Tutor"=ST57Q03,
  "ST57Q04: Out-of-School Study Time - Commercial Company"=ST57Q04,
  "ST57Q05: Out-of-School Study Time - With Parent"=ST57Q05,
  "ST57Q06: Out-of-School Study Time - Computer"=ST57Q06,
  "ST61Q01: Experience with Applied Maths Tasks - Use Train Timetable"=ST61Q01,
  "ST61Q02: Experience with Applied Maths Tasks - Calculate Price including Tax"=ST61Q02,
  "ST61Q03: Experience with Applied Maths Tasks - Calculate Square Metres"=ST61Q03,
  "ST61Q04: Experience with Applied Maths Tasks - Understand Scientific Tables"=ST61Q04,
  "ST61Q05: Experience with Pure Maths Tasks - Solve Equation 1"=ST61Q05,
  "ST61Q06: Experience with Applied Maths Tasks - Use a Map to Calculate Distance"=ST61Q06,
  "ST61Q07: Experience with Pure Maths Tasks - Solve Equation 2"=ST61Q07,
  "ST61Q08: Experience with Applied Maths Tasks - Calculate Power Consumption Rate"=ST61Q08,
  "ST61Q09: Experience with Applied Maths Tasks - Solve Equation 3"=ST61Q09,
  "ST62Q01: Familiarity with Math Concepts - Exponential Function"=ST62Q01,
  "ST62Q02: Familiarity with Math Concepts - Divisor"=ST62Q02,
  "ST62Q03: Familiarity with Math Concepts - Quadratic Function"=ST62Q03,
  "ST62Q04: Overclaiming - Proper Number"=ST62Q04,
  "ST62Q06: Familiarity with Math Concepts - Linear Equation"=ST62Q06,
  "ST62Q07: Familiarity with Math Concepts - Vectors"=ST62Q07,
  "ST62Q08: Familiarity with Math Concepts - Complex Number"=ST62Q08,
  "ST62Q09: Familiarity with Math Concepts - Rational Number"=ST62Q09,
  "ST62Q10: Familiarity with Math Concepts - Radicals"=ST62Q10,
  "ST62Q11: Overclaiming - Subjunctive Scaling"=ST62Q11,
  "ST62Q12: Familiarity with Math Concepts - Polygon"=ST62Q12,
  "ST62Q13: Overclaiming - Declarative Fraction"=ST62Q13,
  "ST62Q15: Familiarity with Math Concepts - Congruent Figure"=ST62Q15,
  "ST62Q16: Familiarity with Math Concepts - Cosine"=ST62Q16,
  "ST62Q17: Familiarity with Math Concepts - Arithmetic Mean"=ST62Q17,
  "ST62Q19: Familiarity with Math Concepts - Probability"=ST62Q19,
  "ST69Q01: Min in <class period> - <test lang>"=ST69Q01,
  "ST69Q02: Min in <class period> - <Maths>"=ST69Q02,
  "ST69Q03: Min in <class period> - <Science>"=ST69Q03,
  "ST70Q01: No of <class period> p/wk - <test lang>"=ST70Q01,
  "ST70Q02: No of <class period> p/wk - <Maths>"=ST70Q02,
  "ST70Q03: No of <class period> p/wk - <Science>"=ST70Q03,
  "ST71Q01: No of ALL <class period> a week"=ST71Q01,
  "ST72Q01: Class Size - No of Students in Test Language Class"=ST72Q01,
  "ST73Q01: OTL - Algebraic Word Problem in Math Lesson"=ST73Q01,
  "ST73Q02: OTL - Algebraic Word Problem in Tests"=ST73Q02,
  "ST74Q01: OTL - Procedural Task in Math Lesson"=ST74Q01,
  "ST74Q02: OTL - Procedural Task in Tests"=ST74Q02,
  "ST75Q01: OTL - Pure Math Reasoning in Math Lesson"=ST75Q01,
  "ST75Q02: OTL - Pure Math Reasoning in Tests"=ST75Q02,
  "ST76Q01: OTL - Applied Math Reasoning in Math Lesson"=ST76Q01,
  "ST76Q02: OTL - Applied Math Reasoning in Tests"=ST76Q02,
  "ST77Q01: Math Teaching - Teacher shows interest"=ST77Q01,
  "ST77Q02: Math Teaching - Extra help"=ST77Q02,
  "ST77Q04: Math Teaching - Teacher helps"=ST77Q04,
  "ST77Q05: Math Teaching - Teacher continues"=ST77Q05,
  "ST77Q06: Math Teaching - Express opinions"=ST77Q06,
  "ST79Q01: Teacher-Directed Instruction - Sets Clear Goals"=ST79Q01,
  "ST79Q02: Teacher-Directed Instruction - Encourages Thinking and Reasoning"=ST79Q02,
  "ST79Q03: Student Orientation - Differentiates Between Students When Giving Tasks"=ST79Q03,
  "ST79Q04: Student Orientation - Assigns Complex Projects"=ST79Q04,
  "ST79Q05: Formative Assessment - Gives Feedback"=ST79Q05,
  "ST79Q06: Teacher-Directed Instruction - Checks Understanding"=ST79Q06,
  "ST79Q07: Student Orientation - Has Students Work in Small Groups"=ST79Q07,
  "ST79Q08: Teacher-Directed Instruction - Summarizes Previous Lessons"=ST79Q08,
  "ST79Q10: Student Orientation - Plans Classroom Activities"=ST79Q10,
  "ST79Q11: Formative Assessment - Gives Feedback on Strengths and Weaknesses"=ST79Q11,
  "ST79Q12: Formative Assessment - Informs about Expectations"=ST79Q12,
  "ST79Q15: Teacher-Directed Instruction - Informs about Learning Goals"=ST79Q15,
  "ST79Q17: Formative Assessment - Tells How to Get Better"=ST79Q17,
  "ST80Q01: Cognitive Activation - Teacher Encourages to Reflect Problems"=ST80Q01,
  "ST80Q04: Cognitive Activation - Gives Problems that Require to Think"=ST80Q04,
  "ST80Q05: Cognitive Activation - Asks to Use Own Procedures"=ST80Q05,
  "ST80Q06: Cognitive Activation - Presents Problems with No Obvious Solutions"=ST80Q06,
  "ST80Q07: Cognitive Activation - Presents Problems in Different Contexts"=ST80Q07,
  "ST80Q08: Cognitive Activation - Helps Learn from Mistakes"=ST80Q08,
  "ST80Q09: Cognitive Activation - Asks for Explanations"=ST80Q09,
  "ST80Q10: Cognitive Activation - Apply What We Learned"=ST80Q10,
  "ST80Q11: Cognitive Activation - Problems with Multiple Solutions"=ST80Q11,
  "ST81Q01: Disciplinary Climate - Students Don\'t Listen"=ST81Q01,
  "ST81Q02: Disciplinary Climate - Noise and Disorder"=ST81Q02,
  "ST81Q03: Disciplinary Climate - Teacher Has to Wait Until its Quiet"=ST81Q03,
  "ST81Q04: Disciplinary Climate - Students Don\'t Work Well"=ST81Q04,
  "ST81Q05: Disciplinary Climate - Students Start Working Late"=ST81Q05,
  "ST82Q01: Vignette Teacher Support - Homework Every Other Day/Back in Time"=ST82Q01,
  "ST82Q02: Vignette Teacher Support - Homework Once a Week/Back in Time"=ST82Q02,
  "ST82Q03: Vignette Teacher Support - Homework Once a Week/Not Back in Time"=ST82Q03,
  "ST83Q01: Teacher Support - Lets Us Know We Have to Work Hard"=ST83Q01,
  "ST83Q02: Teacher Support - Provides Extra Help When Needed"=ST83Q02,
  "ST83Q03: Teacher Support - Helps Students with Learning"=ST83Q03,
  "ST83Q04: Teacher Support - Gives Opportunity to Express Opinions"=ST83Q04,
  "ST84Q01: Vignette Classroom Management - Students Frequently Interrupt/Teacher Arrives Early"=ST84Q01,
  "ST84Q02: Vignette Classroom Management - Students Are Calm/Teacher Arrives on Time"=ST84Q02,
  "ST84Q03: Vignette Classroom Management - Students Frequently Interrupt/Teacher Arrives Late"=ST84Q03,
  "ST85Q01: Classroom Management - Students Listen"=ST85Q01,
  "ST85Q02: Classroom Management - Teacher Keeps Class Orderly"=ST85Q02,
  "ST85Q03: Classroom Management - Teacher Starts On Time"=ST85Q03,
  "ST85Q04: Classroom Management - Wait Long to <Quiet Down>"=ST85Q04,
  "ST86Q01: Student-Teacher Relation - Get Along with Teachers"=ST86Q01,
  "ST86Q02: Student-Teacher Relation - Teachers Are Interested"=ST86Q02,
  "ST86Q03: Student-Teacher Relation - Teachers Listen to Students"=ST86Q03,
  "ST86Q04: Student-Teacher Relation - Teachers Help Students"=ST86Q04,
  "ST86Q05: Student-Teacher Relation - Teachers Treat Students Fair"=ST86Q05,
  "ST87Q01: Sense of Belonging - Feel Like Outsider"=ST87Q01,
  "ST87Q02: Sense of Belonging - Make Friends Easily"=ST87Q02,
  "ST87Q03: Sense of Belonging - Belong at School"=ST87Q03,
  "ST87Q04: Sense of Belonging - Feel Awkward at School"=ST87Q04,
  "ST87Q05: Sense of Belonging - Liked by Other Students"=ST87Q05,
  "ST87Q06: Sense of Belonging - Feel Lonely at School"=ST87Q06,
  "ST87Q07: Sense of Belonging - Feel Happy at School"=ST87Q07,
  "ST87Q08: Sense of Belonging - Things Are Ideal at School"=ST87Q08,
  "ST87Q09: Sense of Belonging - Satisfied at School"=ST87Q09,
  "ST88Q01: Attitude towards School - Does Little to Prepare Me for Life"=ST88Q01,
  "ST88Q02: Attitude towards School - Waste of Time"=ST88Q02,
  "ST88Q03: Attitude towards School - Gave Me Confidence"=ST88Q03,
  "ST88Q04: Attitude towards School - Useful for Job"=ST88Q04,
  "ST89Q02: Attitude toward School - Helps to Get a Job"=ST89Q02,
  "ST89Q03: Attitude toward School - Prepare for College"=ST89Q03,
  "ST89Q04: Attitude toward School - Enjoy Good Grades"=ST89Q04,
  "ST89Q05: Attitude toward School - Trying Hard is Important"=ST89Q05,
  "ST91Q01: Perceived Control - Can Succeed with Enough Effort"=ST91Q01,
  "ST91Q02: Perceived Control - My Choice Whether I Will Be Good"=ST91Q02,
  "ST91Q03: Perceived Control - Problems Prevent from Putting Effort into School"=ST91Q03,
  "ST91Q04: Perceived Control - Different Teachers Would Make Me Try Harder"=ST91Q04,
  "ST91Q05: Perceived Control - Could Perform Well if I Wanted"=ST91Q05,
  "ST91Q06: Perceived Control - Perform Poor Regardless"=ST91Q06,
  "ST93Q01: Perseverance - Give up easily"=ST93Q01,
  "ST93Q03: Perseverance - Put off difficult problems"=ST93Q03,
  "ST93Q04: Perseverance - Remain interested"=ST93Q04,
  "ST93Q06: Perseverance - Continue to perfection"=ST93Q06,
  "ST93Q07: Perseverance - Exceed expectations"=ST93Q07,
  "ST94Q05: Openness for Problem Solving - Can Handle a Lot of Information"=ST94Q05,
  "ST94Q06: Openness for Problem Solving - Quick to Understand"=ST94Q06,
  "ST94Q09: Openness for Problem Solving - Seek Explanations"=ST94Q09,
  "ST94Q10: Openness for Problem Solving - Can Link Facts"=ST94Q10,
  "ST94Q14: Openness for Problem Solving - Like to Solve Complex Problems"=ST94Q14,
  "ST96Q01: Problem Text Message - Press every button"=ST96Q01,
  "ST96Q02: Problem Text Message - Trace steps"=ST96Q02,
  "ST96Q03: Problem Text Message - Manual"=ST96Q03,
  "ST96Q05: Problem Text Message - Ask a friend"=ST96Q05,
  "ST101Q01: Problem Route Selection - Read brochure"=ST101Q01,
  "ST101Q02: Problem Route Selection - Study map"=ST101Q02,
  "ST101Q03: Problem Route Selection - Leave it to brother"=ST101Q03,
  "ST101Q05: Problem Route Selection - Just drive"=ST101Q05,
  "ST104Q01: Problem Ticket Machine - Similarities"=ST104Q01,
  "ST104Q04: Problem Ticket Machine - Try buttons"=ST104Q04,
  "ST104Q05: Problem Ticket Machine - Ask for help"=ST104Q05,
  "ST104Q06: Problem Ticket Machine - Find ticket office"=ST104Q06,
  "IC01Q01: At Home - Desktop Computer"=IC01Q01,
  "IC01Q02: At Home - Portable laptop"=IC01Q02,
  "IC01Q03: At Home - Tablet computer"=IC01Q03,
  "IC01Q04: At Home - Internet connection"=IC01Q04,
  "IC01Q05: At Home - Video games console"=IC01Q05,
  "IC01Q06: At Home - Cell phone w/o Internet"=IC01Q06,
  "IC01Q07: At Home - Cell phone with Internet"=IC01Q07,
  "IC01Q08: At Home - Mp3/Mp4 player"=IC01Q08,
  "IC01Q09: At Home - Printer"=IC01Q09,
  "IC01Q10: At Home - USB (memory) stick"=IC01Q10,
  "IC01Q11: At Home - Ebook reader"=IC01Q11,
  "IC02Q01: At school - Desktop Computer"=IC02Q01,
  "IC02Q02: At school - Portable laptop"=IC02Q02,
  "IC02Q03: At school - Tablet computer"=IC02Q03,
  "IC02Q04: At school - Internet connection"=IC02Q04,
  "IC02Q05: At school - Printer"=IC02Q05,
  "IC02Q06: At school - USB (memory) stick"=IC02Q06,
  "IC02Q07: At school - Ebook reader"=IC02Q07,
  "IC03Q01: First use of computers"=IC03Q01,
  "IC04Q01: First access to Internet"=IC04Q01,
  "IC05Q01: Internet at School"=IC05Q01,
  "IC06Q01: Internet out-of-school - Weekday"=IC06Q01,
  "IC07Q01: Internet out-of-school - Weekend"=IC07Q01,
  "IC08Q01: Out-of-school 8 - One player games."=IC08Q01,
  "IC08Q02: Out-of-school 8 - Collabourative games."=IC08Q02,
  "IC08Q03: Out-of-school 8 - Use email"=IC08Q03,
  "IC08Q04: Out-of-school 8 - Chat on line"=IC08Q04,
  "IC08Q05: Out-of-school 8 - Social networks"=IC08Q05,
  "IC08Q06: Out-of-school 8 - Browse the Internet for fun"=IC08Q06,
  "IC08Q07: Out-of-school 8 - Read news"=IC08Q07,
  "IC08Q08: Out-of-school 8 - Obtain practical information from the Internet"=IC08Q08,
  "IC08Q09: Out-of-school 8 - Download music"=IC08Q09,
  "IC08Q11: Out-of-school 8 - Upload content"=IC08Q11,
  "IC09Q01: Out-of-school 9 - Internet for school"=IC09Q01,
  "IC09Q02: Out-of-school 9 - Email students"=IC09Q02,
  "IC09Q03: Out-of-school 9 - Email teachers"=IC09Q03,
  "IC09Q04: Out-of-school 9 - Download from School"=IC09Q04,
  "IC09Q05: Out-of-school 9 - Announcements"=IC09Q05,
  "IC09Q06: Out-of-school 9 - Homework"=IC09Q06,
  "IC09Q07: Out-of-school 9 - Share school material"=IC09Q07,
  "IC10Q01: At School - Chat on line"=IC10Q01,
  "IC10Q02: At School - Email"=IC10Q02,
  "IC10Q03: At School - Browse for schoolwork"=IC10Q03,
  "IC10Q04: At School - Download from website"=IC10Q04,
  "IC10Q05: At School - Post on website"=IC10Q05,
  "IC10Q06: At School - Simulations"=IC10Q06,
  "IC10Q07: At School - Practice and drilling"=IC10Q07,
  "IC10Q08: At School - Homework"=IC10Q08,
  "IC10Q09: At School - Group work"=IC10Q09,
  "IC11Q01: Maths lessons - Draw graph"=IC11Q01,
  "IC11Q02: Maths lessons - Calculation with numbers"=IC11Q02,
  "IC11Q03: Maths lessons - Geometric figures"=IC11Q03,
  "IC11Q04: Maths lessons - Spreadsheet"=IC11Q04,
  "IC11Q05: Maths lessons - Algebra"=IC11Q05,
  "IC11Q06: Maths lessons - Histograms"=IC11Q06,
  "IC11Q07: Maths lessons - Change in graphs"=IC11Q07,
  "IC22Q01: Attitudes - Useful for schoolwork"=IC22Q01,
  "IC22Q02: Attitudes - Homework more fun"=IC22Q02,
  "IC22Q04: Attitudes - Source of information"=IC22Q04,
  "IC22Q06: Attitudes - Troublesome"=IC22Q06,
  "IC22Q07: Attitudes - Not suitable for schoolwork"=IC22Q07,
  "IC22Q08: Attitudes - Too unreliable"=IC22Q08,
  "EC01Q01: Miss 2 months of <ISCED 1>"=EC01Q01,
  "EC02Q01: Miss 2 months of <ISCED 2>"=EC02Q01,
  "EC03Q01: Future Orientation - Internship"=EC03Q01,
  "EC03Q02: Future Orientation - Work-site visits"=EC03Q02,
  "EC03Q03: Future Orientation - Job fair"=EC03Q03,
  "EC03Q04: Future Orientation - Career advisor at school"=EC03Q04,
  "EC03Q05: Future Orientation - Career advisor outside school"=EC03Q05,
  "EC03Q06: Future Orientation - Questionnaire"=EC03Q06,
  "EC03Q07: Future Orientation - Internet search"=EC03Q07,
  "EC03Q08: Future Orientation - Tour <ISCED 3-5> institution"=EC03Q08,
  "EC03Q09: Future Orientation - web search <ISCED 3-5> prog"=EC03Q09,
  "EC03Q10: Future Orientation - <country specific item>"=EC03Q10,
  "EC04Q01A: Acquired skills - Find job info - Yes, at school"=EC04Q01A,
  "EC04Q01B: Acquired skills - Find job info - Yes, out of school"=EC04Q01B,
  "EC04Q01C: Acquired skills - Find job info - No, never"=EC04Q01C,
  "EC04Q02A: Acquired skills - Search for job - Yes, at school"=EC04Q02A,
  "EC04Q02B: Acquired skills - Search for job - Yes, out of school"=EC04Q02B,
  "EC04Q02C: Acquired skills - Search for job - No, never"=EC04Q02C,
  "EC04Q03A: Acquired skills - Write resume - Yes, at school"=EC04Q03A,
  "EC04Q03B: Acquired skills - Write resume - Yes, out of school"=EC04Q03B,
  "EC04Q03C: Acquired skills - Write resume - No, never"=EC04Q03C,
  "EC04Q04A: Acquired skills - Job interview - Yes, at school"=EC04Q04A,
  "EC04Q04B: Acquired skills - Job interview - Yes, out of school"=EC04Q04B,
  "EC04Q04C: Acquired skills - Job interview - No, never"=EC04Q04C,
  "EC04Q05A: Acquired skills - ISCED 3-5 programs - Yes, at school"=EC04Q05A,
  "EC04Q05B: Acquired skills - ISCED 3-5 programs - Yes, out of school"=EC04Q05B,
  "EC04Q05C: Acquired skills - ISCED 3-5 programs - No, never"=EC04Q05C,
  "EC04Q06A: Acquired skills - Student financing - Yes, at school"=EC04Q06A,
  "EC04Q06B: Acquired skills - Student financing - Yes, out of school"=EC04Q06B,
  "EC04Q06C: Acquired skills - Student financing - No, never"=EC04Q06C,
  "EC05Q01: First language learned"=EC05Q01,
  "EC06Q01: Age started learning test language"=EC06Q01,
  "EC07Q01: Language spoken - Mother"=EC07Q01,
  "EC07Q02: Language spoken - Father"=EC07Q02,
  "EC07Q03: Language spoken - Siblings"=EC07Q03,
  "EC07Q04: Language spoken - Best friend"=EC07Q04,
  "EC07Q05: Language spoken - Schoolmates"=EC07Q05,
  "EC08Q01: Activities language - Reading"=EC08Q01,
  "EC08Q02: Activities language - Watching TV"=EC08Q02,
  "EC08Q03: Activities language - Internet surfing"=EC08Q03,
  "EC08Q04: Activities language - Writing emails"=EC08Q04,
  "EC09Q03: Types of support <test language> - remedial lessons"=EC09Q03,
  "EC10Q01: Amount of support <test language>"=EC10Q01,
  "EC11Q02: Attend lessons <heritage language> - focused"=EC11Q02,
  "EC11Q03: Attend lessons <heritage language> - school subjects"=EC11Q03,
  "EC12Q01: Instruction in <heritage language>"=EC12Q01,
  "ST22Q01: Acculturation - Mother Immigrant (Filter)"=ST22Q01,
  "ST23Q01: Acculturation - Enjoy <Host Culture> Friends"=ST23Q01,
  "ST23Q02: Acculturation - Enjoy <Heritage Culture> Friends"=ST23Q02,
  "ST23Q03: Acculturation - Enjoy <Host Culture> Celebrations"=ST23Q03,
  "ST23Q04: Acculturation - Enjoy <Heritage Culture> Celebrations"=ST23Q04,
  "ST23Q05: Acculturation - Spend Time with <Host Culture> Friends"=ST23Q05,
  "ST23Q06: Acculturation - Spend Time with <Heritage Culture> Friends"=ST23Q06,
  "ST23Q07: Acculturation - Participate in <Host Culture> Celebrations"=ST23Q07,
  "ST23Q08: Acculturation - Participate in <Heritage Culture> Celebrations"=ST23Q08,
  "ST24Q01: Acculturation - Perceived Host-Heritage Cultural Differences - Values"=ST24Q01,
  "ST24Q02: Acculturation - Perceived Host-Heritage Cultural Differences - Mother Treatment"=ST24Q02,
  "ST24Q03: Acculturation - Perceived Host-Heritage Cultural Differences - Teacher Treatment"=ST24Q03,
  "CLCUSE1: Calculator Use"=CLCUSE1,
  "CLCUSE301: Effort-real 1"=CLCUSE301,
  "CLCUSE302: Effort-real 2"=CLCUSE302,
  "DEFFORT: Difference in Effort"=DEFFORT,
  "QUESTID: Student Questionnaire Form"=QUESTID,
  "BOOKID: Booklet ID"=BOOKID,
  "EASY: Standard or simplified set of booklets"=EASY,
  "AGE: Age of student"=AGE,
  "GRADE: Grade compared to modal grade in country"=GRADE,
  "PROGN: Unique national study programme code"=PROGN,
  "ANXMAT: Mathematics Anxiety"=ANXMAT,
  "ATSCHL: Attitude towards School: Learning Outcomes"=ATSCHL,
  "ATTLNACT: Attitude towards School: Learning Activities"=ATTLNACT,
  "BELONG: Sense of Belonging to School"=BELONG,
  "BFMJ2: Father SQ ISEI"=BFMJ2,
  "BMMJ1: Mother SQ ISEI"=BMMJ1,
  "CLSMAN: Mathematics Teacher's Classroom Management"=CLSMAN,
  "COBN_F: Country of Birth National Categories - Father"=COBN_F,
  "COBN_M: Country of Birth National Categories - Mother"=COBN_M,
  "COBN_S: Country of Birth National Categories - Self"=COBN_S,
  "COGACT: Cognitive Activation in Mathematics Lessons"=COGACT,
  "CULTDIST: Cultural Distance between Host and Heritage Culture"=CULTDIST,
  "CULTPOS: Cultural Possessions"=CULTPOS,
  "DISCLIMA: Disciplinary Climate"=DISCLIMA,
  "ENTUSE: ICT Entertainment Use"=ENTUSE,
  "ESCS: Index of economic, social and cultural status"=ESCS,
  "EXAPPLM: Experience with Applied Mathematics Tasks at School"=EXAPPLM,
  "EXPUREM: Experience with Pure Mathematics Tasks at School"=EXPUREM,
  "FAILMAT: Attributions to Failure in Mathematics"=FAILMAT,
  "FAMCON: Familiarity with Mathematical Concepts"=FAMCON,
  "FAMCONC: Familiarity with Mathematical Concepts (Signal Detection Adjusted)"=FAMCONC,
  "FAMSTRUC: Family Structure"=FAMSTRUC,
  "FISCED: Educational level of father (ISCED)"=FISCED,
  "HEDRES: Home educational resources"=HEDRES,
  "HERITCUL: Acculturation: Heritage Culture Oriented Strategies"=HERITCUL,
  "HISCED: Highest educational level of parents"=HISCED,
  "HISEI: Highest parental occupational status"=HISEI,
  "HOMEPOS: Home Possessions"=HOMEPOS,
  "HOMSCH: ICT Use at Home for School-related Tasks"=HOMSCH,
  "HOSTCUL: Acculturation: Host Culture Oriented Strategies"=HOSTCUL,
  "ICTATTNEG: Attitudes Towards Computers: Limitations of the Computer as a Tool for School Learning"=ICTATTNEG,
  "ICTATTPOS: Attitudes Towards Computers: Computer as a Tool for School Learning"=ICTATTPOS,
  "ICTHOME: ICT Availability at Home"=ICTHOME,
  "ICTRES: ICT resources"=ICTRES,
  "ICTSCH: ICT Availability at School"=ICTSCH,
  "IMMIG: Immigration status"=IMMIG,
  "INFOCAR: Information about Careers"=INFOCAR,
  "INFOJOB1: Information about the Labour Market provided by the School"=INFOJOB1,
  "INFOJOB2: Information about the Labour Market provided outside of School"=INFOJOB2,
  "INSTMOT: Instrumental Motivation for Mathematics"=INSTMOT,
  "INTMAT: Mathematics Interest"=INTMAT,
  "ISCEDD: ISCED designation"=ISCEDD,
  "ISCEDL: ISCED level"=ISCEDL,
  "ISCEDO: ISCED orientation"=ISCEDO,
  "LANGCOMM: Preference for Heritage Language in Conversations with Family and Friends"=LANGCOMM,
  "LANGN: Language at home (3-digit code)"=LANGN,
  "LANGRPPD: Preference for Heritage Language in Language Reception and Production"=LANGRPPD,
  "LMINS: Learning time (minutes per week) - <test language>"=LMINS,
  "MATBEH: Mathematics Behaviour"=MATBEH,
  "MATHEFF: Mathematics Self-Efficacy"=MATHEFF,
  "MATINTFC: Mathematics Intentions"=MATINTFC,
  "MATWKETH: Mathematics Work Ethic"=MATWKETH,
  "MISCED: Educational level of mother (ISCED)"=MISCED,
  "MMINS: Learning time (minutes per week) - <Mathematics>"=MMINS,
  "MTSUP: Mathematics Teacher's Support"=MTSUP,
  "OCOD1: ISCO-08 Occupation code - Mother"=OCOD1,
  "OCOD2: ISCO-08 Occupation code - Father"=OCOD2,
  "OPENPS: Openness for Problem Solving"=OPENPS,
  "OUTHOURS: Out-of-School Study Time"=OUTHOURS,
  "PARED: Highest parental education in years"=PARED,
  "PERSEV: Perseverance"=PERSEV,
  "REPEAT: Grade Repetition"=REPEAT,
  "SCMAT: Mathematics Self-Concept"=SCMAT,
  "SMINS: Learning time (minutes per week) - <Science>"=SMINS,
  "STUDREL: Teacher Student Relations"=STUDREL,
  "SUBNORM: Subjective Norms in Mathematics"=SUBNORM,
  "TCHBEHFA: Teacher Behaviour: Formative Assessment"=TCHBEHFA,
  "TCHBEHSO: Teacher Behaviour: Student Orientation"=TCHBEHSO,
  "TCHBEHTD: Teacher Behaviour: Teacher-directed Instruction"=TCHBEHTD,
  "TEACHSUP: Teacher Support"=TEACHSUP,
  "TESTLANG: Language of the test"=TESTLANG,
  "TIMEINT: Time of computer use (mins)"=TIMEINT,
  "USEMATH: Use of ICT in Mathematic Lessons"=USEMATH,
  "USESCH: Use of ICT at School"=USESCH,
  "WEALTH: Wealth"=WEALTH,
  "ANCATSCHL: Attitude towards School: Learning Outcomes (Anchored)"=ANCATSCHL,
  "ANCATTLNACT Attitude towards School: Learning Activities (Anchored)"=ANCATTLNACT,
  "ANCBELONG: Sense of Belonging to School (Anchored)"=ANCBELONG,
  "ANCCLSMAN: Mathematics Teacher's Classroom Management (Anchored)"=ANCCLSMAN,
  "ANCCOGACT: Cognitive Activation in Mathematics Lessons (Anchored)"=ANCCOGACT,
  "ANCINSTMOT: Instrumental Motivation for Mathematics (Anchored)"=ANCINSTMOT,
  "ANCINTMAT: Mathematics Interest (Anchored)"=ANCINTMAT,
  "ANCMATWKETH Mathematics Work Ethic (Anchored)"=ANCMATWKETH,
  "ANCMTSUP: Mathematics Teacher's Support (Anchored)"=ANCMTSUP,
  "ANCSCMAT: Mathematics Self-Concept (Anchored)"=ANCSCMAT,
  "ANCSTUDREL: Teacher Student Relations (Anchored)"=ANCSTUDREL,
  "ANCSUBNORM: Subjective Norms in Mathematics (Anchored)"=ANCSUBNORM,
  "PV1MATH: Plausible value 1 in mathematics"=PV1MATH,
  "PV2MATH: Plausible value 2 in mathematics"=PV2MATH,
  "PV3MATH: Plausible value 3 in mathematics"=PV3MATH,
  "PV4MATH: Plausible value 4 in mathematics"=PV4MATH,
  "PV5MATH: Plausible value 5 in mathematics"=PV5MATH,
  "PV1MACC: Plausible value 1 in content subscale of math - Change and Relationships"=PV1MACC,
  "PV2MACC: Plausible value 2 in content subscale of math - Change and Relationships"=PV2MACC,
  "PV3MACC: Plausible value 3 in content subscale of math - Change and Relationships"=PV3MACC,
  "PV4MACC: Plausible value 4 in content subscale of math - Change and Relationships"=PV4MACC,
  "PV5MACC: Plausible value 5 in content subscale of math - Change and Relationships"=PV5MACC,
  "PV1MACQ: Plausible value 1 in content subscale of math - Quantity"=PV1MACQ,
  "PV2MACQ: Plausible value 2 in content subscale of math - Quantity"=PV2MACQ,
  "PV3MACQ: Plausible value 3 in content subscale of math - Quantity"=PV3MACQ,
  "PV4MACQ: Plausible value 4 in content subscale of math - Quantity"=PV4MACQ,
  "PV5MACQ: Plausible value 5 in content subscale of math - Quantity"=PV5MACQ,
  "PV1MACS: Plausible value 1 in content subscale of math - Space and Shape"=PV1MACS,
  "PV2MACS: Plausible value 2 in content subscale of math - Space and Shape"=PV2MACS,
  "PV3MACS: Plausible value 3 in content subscale of math - Space and Shape"=PV3MACS,
  "PV4MACS: Plausible value 4 in content subscale of math - Space and Shape"=PV4MACS,
  "PV5MACS: Plausible value 5 in content subscale of math - Space and Shape"=PV5MACS,
  "PV1MACU: Plausible value 1 in content subscale of math - Uncertainty and Data"=PV1MACU,
  "PV2MACU: Plausible value 2 in content subscale of math - Uncertainty and Data"=PV2MACU,
  "PV3MACU: Plausible value 3 in content subscale of math - Uncertainty and Data"=PV3MACU,
  "PV4MACU: Plausible value 4 in content subscale of math - Uncertainty and Data"=PV4MACU,
  "PV5MACU: Plausible value 5 in content subscale of math - Uncertainty and Data"=PV5MACU,
  "PV1MAPE: Plausible value 1 in process subscale of math - Employ"=PV1MAPE,
  "PV2MAPE: Plausible value 2 in process subscale of math - Employ"=PV2MAPE,
  "PV3MAPE: Plausible value 3 in process subscale of math - Employ"=PV3MAPE,
  "PV4MAPE: Plausible value 4 in process subscale of math - Employ"=PV4MAPE,
  "PV5MAPE: Plausible value 5 in process subscale of math - Employ"=PV5MAPE,
  "PV1MAPF: Plausible value 1 in process subscale of math - Formulate"=PV1MAPF,
  "PV2MAPF: Plausible value 2 in process subscale of math - Formulate"=PV2MAPF,
  "PV3MAPF: Plausible value 3 in process subscale of math - Formulate"=PV3MAPF,
  "PV4MAPF: Plausible value 4 in process subscale of math - Formulate"=PV4MAPF,
  "PV5MAPF: Plausible value 5 in process subscale of math - Formulate"=PV5MAPF,
  "PV1MAPI: Plausible value 1 in process subscale of math - Interpret"=PV1MAPI,
  "PV2MAPI: Plausible value 2 in process subscale of math - Interpret"=PV2MAPI,
  "PV3MAPI: Plausible value 3 in process subscale of math - Interpret"=PV3MAPI,
  "PV4MAPI: Plausible value 4 in process subscale of math - Interpret"=PV4MAPI,
  "PV5MAPI: Plausible value 5 in process subscale of math - Interpret"=PV5MAPI,
  "PV1READ: Plausible value 1 in reading"=PV1READ,
  "PV2READ: Plausible value 2 in reading"=PV2READ,
  "PV3READ: Plausible value 3 in reading"=PV3READ,
  "PV4READ: Plausible value 4 in reading"=PV4READ,
  "PV5READ: Plausible value 5 in reading"=PV5READ,
  "PV1SCIE: Plausible value 1 in science"=PV1SCIE,
  "PV2SCIE: Plausible value 2 in science"=PV2SCIE,
  "PV3SCIE: Plausible value 3 in science"=PV3SCIE,
  "PV4SCIE: Plausible value 4 in science"=PV4SCIE,
  "PV5SCIE: Plausible value 5 in science"=PV5SCIE,
  "W_FSTUWT: FINAL STUDENT WEIGHT"=W_FSTUWT,
  "W_FSTR1: FINAL STUDENT REPLICATE BRR-FAY WEIGHT1"=W_FSTR1,
  "W_FSTR2: FINAL STUDENT REPLICATE BRR-FAY WEIGHT2"=W_FSTR2,
  "W_FSTR3: FINAL STUDENT REPLICATE BRR-FAY WEIGHT3"=W_FSTR3,
  "W_FSTR4: FINAL STUDENT REPLICATE BRR-FAY WEIGHT4"=W_FSTR4,
  "W_FSTR5: FINAL STUDENT REPLICATE BRR-FAY WEIGHT5"=W_FSTR5,
  "W_FSTR6: FINAL STUDENT REPLICATE BRR-FAY WEIGHT6"=W_FSTR6,
  "W_FSTR7: FINAL STUDENT REPLICATE BRR-FAY WEIGHT7"=W_FSTR7,
  "W_FSTR8: FINAL STUDENT REPLICATE BRR-FAY WEIGHT8"=W_FSTR8,
  "W_FSTR9: FINAL STUDENT REPLICATE BRR-FAY WEIGHT9"=W_FSTR9,
  "W_FSTR10: FINAL STUDENT REPLICATE BRR-FAY WEIGHT10"=W_FSTR10,
  "W_FSTR11: FINAL STUDENT REPLICATE BRR-FAY WEIGHT11"=W_FSTR11,
  "W_FSTR12: FINAL STUDENT REPLICATE BRR-FAY WEIGHT12"=W_FSTR12,
  "W_FSTR13: FINAL STUDENT REPLICATE BRR-FAY WEIGHT13"=W_FSTR13,
  "W_FSTR14: FINAL STUDENT REPLICATE BRR-FAY WEIGHT14"=W_FSTR14,
  "W_FSTR15: FINAL STUDENT REPLICATE BRR-FAY WEIGHT15"=W_FSTR15,
  "W_FSTR16: FINAL STUDENT REPLICATE BRR-FAY WEIGHT16"=W_FSTR16,
  "W_FSTR17: FINAL STUDENT REPLICATE BRR-FAY WEIGHT17"=W_FSTR17,
  "W_FSTR18: FINAL STUDENT REPLICATE BRR-FAY WEIGHT18"=W_FSTR18,
  "W_FSTR19: FINAL STUDENT REPLICATE BRR-FAY WEIGHT19"=W_FSTR19,
  "W_FSTR20: FINAL STUDENT REPLICATE BRR-FAY WEIGHT20"=W_FSTR20,
  "W_FSTR21: FINAL STUDENT REPLICATE BRR-FAY WEIGHT21"=W_FSTR21,
  "W_FSTR22: FINAL STUDENT REPLICATE BRR-FAY WEIGHT22"=W_FSTR22,
  "W_FSTR23: FINAL STUDENT REPLICATE BRR-FAY WEIGHT23"=W_FSTR23,
  "W_FSTR24: FINAL STUDENT REPLICATE BRR-FAY WEIGHT24"=W_FSTR24,
  "W_FSTR25: FINAL STUDENT REPLICATE BRR-FAY WEIGHT25"=W_FSTR25,
  "W_FSTR26: FINAL STUDENT REPLICATE BRR-FAY WEIGHT26"=W_FSTR26,
  "W_FSTR27: FINAL STUDENT REPLICATE BRR-FAY WEIGHT27"=W_FSTR27,
  "W_FSTR28: FINAL STUDENT REPLICATE BRR-FAY WEIGHT28"=W_FSTR28,
  "W_FSTR29: FINAL STUDENT REPLICATE BRR-FAY WEIGHT29"=W_FSTR29,
  "W_FSTR30: FINAL STUDENT REPLICATE BRR-FAY WEIGHT30"=W_FSTR30,
  "W_FSTR31: FINAL STUDENT REPLICATE BRR-FAY WEIGHT31"=W_FSTR31,
  "W_FSTR32: FINAL STUDENT REPLICATE BRR-FAY WEIGHT32"=W_FSTR32,
  "W_FSTR33: FINAL STUDENT REPLICATE BRR-FAY WEIGHT33"=W_FSTR33,
  "W_FSTR34: FINAL STUDENT REPLICATE BRR-FAY WEIGHT34"=W_FSTR34,
  "W_FSTR35: FINAL STUDENT REPLICATE BRR-FAY WEIGHT35"=W_FSTR35,
  "W_FSTR36: FINAL STUDENT REPLICATE BRR-FAY WEIGHT36"=W_FSTR36,
  "W_FSTR37: FINAL STUDENT REPLICATE BRR-FAY WEIGHT37"=W_FSTR37,
  "W_FSTR38: FINAL STUDENT REPLICATE BRR-FAY WEIGHT38"=W_FSTR38,
  "W_FSTR39: FINAL STUDENT REPLICATE BRR-FAY WEIGHT39"=W_FSTR39,
  "W_FSTR40: FINAL STUDENT REPLICATE BRR-FAY WEIGHT40"=W_FSTR40,
  "W_FSTR41: FINAL STUDENT REPLICATE BRR-FAY WEIGHT41"=W_FSTR41,
  "W_FSTR42: FINAL STUDENT REPLICATE BRR-FAY WEIGHT42"=W_FSTR42,
  "W_FSTR43: FINAL STUDENT REPLICATE BRR-FAY WEIGHT43"=W_FSTR43,
  "W_FSTR44: FINAL STUDENT REPLICATE BRR-FAY WEIGHT44"=W_FSTR44,
  "W_FSTR45: FINAL STUDENT REPLICATE BRR-FAY WEIGHT45"=W_FSTR45,
  "W_FSTR46: FINAL STUDENT REPLICATE BRR-FAY WEIGHT46"=W_FSTR46,
  "W_FSTR47: FINAL STUDENT REPLICATE BRR-FAY WEIGHT47"=W_FSTR47,
  "W_FSTR48: FINAL STUDENT REPLICATE BRR-FAY WEIGHT48"=W_FSTR48,
  "W_FSTR49: FINAL STUDENT REPLICATE BRR-FAY WEIGHT49"=W_FSTR49,
  "W_FSTR50: FINAL STUDENT REPLICATE BRR-FAY WEIGHT50"=W_FSTR50,
  "W_FSTR51: FINAL STUDENT REPLICATE BRR-FAY WEIGHT51"=W_FSTR51,
  "W_FSTR52: FINAL STUDENT REPLICATE BRR-FAY WEIGHT52"=W_FSTR52,
  "W_FSTR53: FINAL STUDENT REPLICATE BRR-FAY WEIGHT53"=W_FSTR53,
  "W_FSTR54: FINAL STUDENT REPLICATE BRR-FAY WEIGHT54"=W_FSTR54,
  "W_FSTR55: FINAL STUDENT REPLICATE BRR-FAY WEIGHT55"=W_FSTR55,
  "W_FSTR56: FINAL STUDENT REPLICATE BRR-FAY WEIGHT56"=W_FSTR56,
  "W_FSTR57: FINAL STUDENT REPLICATE BRR-FAY WEIGHT57"=W_FSTR57,
  "W_FSTR58: FINAL STUDENT REPLICATE BRR-FAY WEIGHT58"=W_FSTR58,
  "W_FSTR59: FINAL STUDENT REPLICATE BRR-FAY WEIGHT59"=W_FSTR59,
  "W_FSTR60: FINAL STUDENT REPLICATE BRR-FAY WEIGHT60"=W_FSTR60,
  "W_FSTR61: FINAL STUDENT REPLICATE BRR-FAY WEIGHT61"=W_FSTR61,
  "W_FSTR62: FINAL STUDENT REPLICATE BRR-FAY WEIGHT62"=W_FSTR62,
  "W_FSTR63: FINAL STUDENT REPLICATE BRR-FAY WEIGHT63"=W_FSTR63,
  "W_FSTR64: FINAL STUDENT REPLICATE BRR-FAY WEIGHT64"=W_FSTR64,
  "W_FSTR65: FINAL STUDENT REPLICATE BRR-FAY WEIGHT65"=W_FSTR65,
  "W_FSTR66: FINAL STUDENT REPLICATE BRR-FAY WEIGHT66"=W_FSTR66,
  "W_FSTR67: FINAL STUDENT REPLICATE BRR-FAY WEIGHT67"=W_FSTR67,
  "W_FSTR68: FINAL STUDENT REPLICATE BRR-FAY WEIGHT68"=W_FSTR68,
  "W_FSTR69: FINAL STUDENT REPLICATE BRR-FAY WEIGHT69"=W_FSTR69,
  "W_FSTR70: FINAL STUDENT REPLICATE BRR-FAY WEIGHT70"=W_FSTR70,
  "W_FSTR71: FINAL STUDENT REPLICATE BRR-FAY WEIGHT71"=W_FSTR71,
  "W_FSTR72: FINAL STUDENT REPLICATE BRR-FAY WEIGHT72"=W_FSTR72,
  "W_FSTR73: FINAL STUDENT REPLICATE BRR-FAY WEIGHT73"=W_FSTR73,
  "W_FSTR74: FINAL STUDENT REPLICATE BRR-FAY WEIGHT74"=W_FSTR74,
  "W_FSTR75: FINAL STUDENT REPLICATE BRR-FAY WEIGHT75"=W_FSTR75,
  "W_FSTR76: FINAL STUDENT REPLICATE BRR-FAY WEIGHT76"=W_FSTR76,
  "W_FSTR77: FINAL STUDENT REPLICATE BRR-FAY WEIGHT77"=W_FSTR77,
  "W_FSTR78: FINAL STUDENT REPLICATE BRR-FAY WEIGHT78"=W_FSTR78,
  "W_FSTR79: FINAL STUDENT REPLICATE BRR-FAY WEIGHT79"=W_FSTR79,
  "W_FSTR80: FINAL STUDENT REPLICATE BRR-FAY WEIGHT80"=W_FSTR80,
  "WVARSTRR: RANDOMIZED FINAL VARIANCE STRATUM (1-80)"=WVARSTRR,
  "VAR_UNIT: RANDOMLY ASSIGNED VARIANCE UNIT"=VAR_UNIT,
  "SENWGT_STU: Senate weight - sum of weight within the country is 1000"=SENWGT_STU)
```

#### Regional data

```{r rename composite columns, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
country_data <- country_data %>%
  rename("CNT: Country code 3-character"=CNT)

subregion_data <- subregion_data %>%
  rename("STRATUM: Stratum ID 7-character (cnt + region ID + original stratum ID)"=STRATUM)
```

### Checking levels

Here I check that the values in all levels look sane.

```{r check levels, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
# summary(pisa[,1:100])
# summary(pisa[,101:200])
# summary(pisa[,201:300])
# summary(pisa[,301:400])
# summary(pisa[,401:500])
# summary(pisa[,501:600])
# summary(pisa[,601:628])
```

### Removing unneeded columns

The row index and date the database was created are not needed.

```{r remove columns, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
remove <- c("V1", "VER_STU")
pisa <- select(pisa, -one_of(remove))
```

## Preparing data for plotting

Now that I've created a few preliminary plots, I have a general idea of what I want to look at, and the extent to which various survey measures (of student/teacher engagement, time spent on learning, etc.) predict success at the country and subregion level.

As I've found that a measure of wealth inequality (not only average wealth) may be a useful measure, I will use the Gini coefficient (https://en.wikipedia.org/wiki/Gini_coefficient) instead of the standard deviation as an estimate for this measure.

For relative ease of interpretation, since some of the measures (e.g., the wealth calculation) are already opaque, I will both average and scale all quantitative measures I might want to plot.

### Country

```{r country data, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
measures_by_country <- pisa %>%
  group_by(`CNT: Country code 3-character`) %>%
  summarize("WEALTH: Wealth - mean (country)" = mean(`WEALTH: Wealth`, na.rm=TRUE), "WEALTH: Wealth - median (country)" = median(`WEALTH: Wealth`, na.rm=TRUE), "WEALTH: Wealth Inequality (country)" = Gini(`WEALTH: Wealth`+10, na.rm=TRUE)*-1, "ESCS: Index of economic, social and cultural status (country)" = mean(`ESCS: Index of economic, social and cultural status`, na.rm=TRUE), "ST57Q01: Out-of-School Study Time - Homework - mean (country)" = mean(`ST57Q01: Out-of-School Study Time - Homework`, na.rm=TRUE), "ST69Q01: Min in <class period> - <test lang> - mean (country)" = mean(`ST69Q01: Min in <class period> - <test lang>`, na.rm=TRUE), "ST69Q02: Min in <class period> - <Maths> - mean (country)" = mean(`ST69Q02: Min in <class period> - <Maths>`, na.rm=TRUE), "ST69Q03: Min in <class period> - <Science> - mean (country)" = mean(`ST69Q03: Min in <class period> - <Science>`, na.rm=TRUE), "IC05Q01: Internet at School - mean (country)" = mean(`IC05Q01: Internet at School`, na.rm=TRUE), "IC06Q01: Internet out-of-school - Weekday - mean (country)" = mean(`IC06Q01: Internet out-of-school - Weekday`, na.rm=TRUE), "IC07Q01: Internet out-of-school - Weekend - mean (country)" = mean(`IC07Q01: Internet out-of-school - Weekend`, na.rm=TRUE), "ATSCHL: Attitude towards School: Learning Outcomes - mean (country)" = mean(`ATSCHL: Attitude towards School: Learning Outcomes`, na.rm=TRUE), "ICTRES: ICT resources - mean (country)" = mean(`ICTRES: ICT resources`, na.rm=TRUE), "INSTMOT: Instrumental Motivation for Mathematics - mean (country)" = mean(`INSTMOT: Instrumental Motivation for Mathematics`, na.rm=TRUE), "INTMAT: Mathematics Interest - mean (country)" = mean(`INTMAT: Mathematics Interest`, na.rm=TRUE), "LMINS: Learning time (minutes per week) - <test language> - mean (country)" = mean(`LMINS: Learning time (minutes per week) - <test language>`, na.rm=TRUE), "MMINS: Learning time (minutes per week) - <Mathematics> - mean (country)" = mean(`MMINS: Learning time (minutes per week) - <Mathematics>`, na.rm=TRUE), "SMINS: Learning time (minutes per week) - <Science> - mean (country)" = mean(`SMINS: Learning time (minutes per week) - <Science>`, na.rm=TRUE), "MATWKETH: Mathematics Work Ethic - mean (country)" = mean(`MATWKETH: Mathematics Work Ethic`, na.rm=TRUE), "PERSEV: Perseverance - mean (country)" = mean(`PERSEV: Perseverance`, na.rm=TRUE), "PARED: Highest parental education in years - mean (country)" = mean(`PARED: Highest parental education in years`, na.rm=TRUE), "OUTHOURS: Out-of-School Study Time - mean (country)" = mean(`OUTHOURS: Out-of-School Study Time`, na.rm=TRUE), "OPENPS: Openness for Problem Solving - mean (country)" = mean(`OPENPS: Openness for Problem Solving`, na.rm=TRUE), "TCHBEHSO: Teacher Behaviour: Student Orientation - mean (country)" = mean(`TCHBEHSO: Teacher Behaviour: Student Orientation`, na.rm=TRUE), "TCHBEHTD: Teacher Behaviour: Teacher-directed Instruction - mean (country)" = mean(`TCHBEHTD: Teacher Behaviour: Teacher-directed Instruction`, na.rm=TRUE), "TEACHSUP: Teacher Support - mean (country)" = mean(`TEACHSUP: Teacher Support`, na.rm=TRUE))

measures_by_country <- left_join(measures_by_country, country_data, by="CNT: Country code 3-character")

for (col in colnames(measures_by_country[,-1])) {
  measures_by_country[[paste(col,"- scaled")]] <- rescale(measures_by_country[[col]], to=c(-0.5,0.5))
}
```

### Region

```{r region data, cache=TRUE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
measures_by_region <- pisa %>%
  group_by(`CNT: Country code 3-character`,`STRATUM: Stratum ID 7-character (cnt + region ID + original stratum ID)`) %>%
  summarize("WEALTH: Wealth - mean (region)" = mean(`WEALTH: Wealth`, na.rm=TRUE), "WEALTH: Wealth - median (region)" = median(`WEALTH: Wealth`, na.rm=TRUE), "WEALTH: Wealth Inequality (region)" = Gini(`WEALTH: Wealth`+10, na.rm=TRUE)*-1, "ESCS: Index of economic, social and cultural status (region)" = mean(`ESCS: Index of economic, social and cultural status`, na.rm=TRUE), "ST57Q01: Out-of-School Study Time - Homework - mean (region)" = mean(`ST57Q01: Out-of-School Study Time - Homework`, na.rm=TRUE), "ST69Q01: Min in <class period> - <test lang> - mean (region)" = mean(`ST69Q01: Min in <class period> - <test lang>`, na.rm=TRUE), "ST69Q02: Min in <class period> - <Maths> - mean (region)" = mean(`ST69Q02: Min in <class period> - <Maths>`, na.rm=TRUE), "ST69Q03: Min in <class period> - <Science> - mean (region)" = mean(`ST69Q03: Min in <class period> - <Science>`, na.rm=TRUE), "IC05Q01: Internet at School - mean (region)" = mean(`IC05Q01: Internet at School`, na.rm=TRUE), "IC06Q01: Internet out-of-school - Weekday - mean (region)" = mean(`IC06Q01: Internet out-of-school - Weekday`, na.rm=TRUE), "IC07Q01: Internet out-of-school - Weekend - mean (region)" = mean(`IC07Q01: Internet out-of-school - Weekend`, na.rm=TRUE), "ATSCHL: Attitude towards School: Learning Outcomes - mean (region)" = mean(`ATSCHL: Attitude towards School: Learning Outcomes`, na.rm=TRUE), "ICTRES: ICT resources - mean (region)" = mean(`ICTRES: ICT resources`, na.rm=TRUE), "INSTMOT: Instrumental Motivation for Mathematics - mean (region)" = mean(`INSTMOT: Instrumental Motivation for Mathematics`, na.rm=TRUE), "INTMAT: Mathematics Interest - mean (region)" = mean(`INTMAT: Mathematics Interest`, na.rm=TRUE), "LMINS: Learning time (minutes per week) - <test language> - mean (region)" = mean(`LMINS: Learning time (minutes per week) - <test language>`, na.rm=TRUE), "MMINS: Learning time (minutes per week) - <Mathematics> - mean (region)" = mean(`MMINS: Learning time (minutes per week) - <Mathematics>`, na.rm=TRUE), "SMINS: Learning time (minutes per week) - <Science> - mean (region)" = mean(`SMINS: Learning time (minutes per week) - <Science>`, na.rm=TRUE), "MATWKETH: Mathematics Work Ethic - mean (region)" = mean(`MATWKETH: Mathematics Work Ethic`, na.rm=TRUE), "PERSEV: Perseverance - mean (region)" = mean(`PERSEV: Perseverance`, na.rm=TRUE), "PARED: Highest parental education in years - mean (region)" = mean(`PARED: Highest parental education in years`, na.rm=TRUE), "OUTHOURS: Out-of-School Study Time - mean (region)" = mean(`OUTHOURS: Out-of-School Study Time`, na.rm=TRUE), "OPENPS: Openness for Problem Solving - mean (region)" = mean(`OPENPS: Openness for Problem Solving`, na.rm=TRUE), "TCHBEHSO: Teacher Behaviour: Student Orientation - mean (region)" = mean(`TCHBEHSO: Teacher Behaviour: Student Orientation`, na.rm=TRUE), "TCHBEHTD: Teacher Behaviour: Teacher-directed Instruction - mean (region)" = mean(`TCHBEHTD: Teacher Behaviour: Teacher-directed Instruction`, na.rm=TRUE), "TEACHSUP: Teacher Support - mean (region)" = mean(`TEACHSUP: Teacher Support`, na.rm=TRUE))

measures_by_region <- left_join(measures_by_region, subregion_data, by=c("STRATUM: Stratum ID 7-character (cnt + region ID + original stratum ID)"))

for (col in colnames(measures_by_region[,-c(1,2)])) {
  measures_by_region[[paste(col,"- scaled")]] <- rescale(measures_by_region[[col]], to=c(-.5,0.5))
}
```

## Exporting data

Save all relevant data (modified PISA dataset and by-country/subregion data) to .csv files in current directory.

```{r export data, message=FALSE, warning=FALSE}
pisa <- select(pisa, -("PV1MATH: Plausible value 1 in mathematics":"W_FSTR80: FINAL STUDENT REPLICATE BRR-FAY WEIGHT80"))

# fwrite(pisa, "pisa.csv")
# fwrite(measures_by_country, "country_data.csv")
# fwrite(measures_by_region, "subregion_data.csv")
```

